Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-01-26 Ben_Manuscripts/stochastic_transport/Final_Draft.tex
Final readthrough completed

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore two variations of the anomalous diffusion model. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies two sets of parameters based on the solute's radial distance from the
    closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which has separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences 
  between MD and Markov state dependent model-generated trajectories may limit 
  its usefulness. Finally, we demonstrate how one can use these models to 
  estimate flux of a solute across a macroscopic-length pore and, based on 
  those quantities, the membrane's selectivity towards each solute. This work 
  helps to connect microscopic chemically-dependent solute motions that do not
  follow simple diffusive behavior with macroscopic membrane performance.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that can be 
  cross-linked into mechanically strong and highly selective nanoporous membranes.~\cite{gin_polymerized_2008}
  They may provide a promising alternative to conventional membrane separation techniques
  by being selective based not only on solute size and charge, but on solute chemical 
  functionality as well.~\cite{dischinger_application_2017} One can tune the functionality
  of LLC monomers in order to enhance or weaken specific solute-membrane interactions.~\cite{dischinger_effect_2017}
  Two general morphologies of LLC membranes are being actively developed. The inverted hexagonal, 
  H\textsubscript{II}, LLC phase has densely packed, uniform-sized pores on the order of 1 nm
  in size. A perfectly synthesized H\textsubscript{II} phase has an ideal geometry for high
  throughput separations. Unfortunately, aligning the microscopic hexagonal mesophases on 
  a large scale as required for high permeability has shown limited success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLCs share the uniform and nm-sized chemically complex pores of
  the H\textsubscript{II} phase but its geometry consists of a tortuous network of 3D 
  interconnected pores which do not require alignment.~\cite{carter_glycerol-based_2012} While 
  its pore structure may decrease a Q\textsubscript{I} membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to synthesize at scale.~\cite{hatakeyama_nanoporous_2010}

@@

  Using a bottom-up modeling approach, we can parameterize single particle behavior
  by extracting their trajectories from MD simulations and studying the properties that
  have the greatest influence on solute dynamics. With this information, we can learn
  how to construct an ensemble of characteristic single solute trajectories with these
  properties which would be useful for making long timescale predictions with computational
  ease and lower uncertainties. Fortunately, there is an abundance of information encoded
  in the complex solute trajectories. We can incorporate fluctuations and time-dependent 
  correlations in solute position as well as the length of time solutes are trapped into a
  model. We can add further detail by integrating our time series analysis with our 
  knowledge of the primary trapping mechanisms as well as the solute's changing chemical
  environment within the heterogeneous membrane.

  %MRS8: really nice flow in the introduction up through here.
%  By reducing our MD simulations to observations of several key properties that 
%  have the greatest influence on solute dynamics, we can construct many realizations
%  of single solute trajectories useful for making long timescale predictions with 
%  computational ease and lower uncertainties. Fortunately, there is an abundance of 
%  information encoded in the complex solute trajectories. Into a model, we can 
%  incorporate fluctuations and time-dependent correlations in solute position as well
%  as the length of time solutes are trapped. We can add further detail by integrating
%  our time series analysis with our knowledge of the primary trapping mechanisms as
%  well as the solute's changing chemical environment within the heterogeneous membrane.

  In this work, we use the output of our MD simulations to construct two classes of 
  mathematical models which aim to predict membrane performance while providing quantitative 
  mechanistic insights. The functional form of these models is driven by mechanistic 
  observations from our previous work and their inputs are parameterized using a 
  substantial amount of data generated by long ($\geq 5~\mu$s) MD simulations.

@@

  Three well-known classes of behavior leading to anomalous subdiffusion are 
  continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  These types of motion are frequently used alone or in combination to describe
  single particle trajectories.~\cite{morrin_three_2018,metzler_anomalous_2014}
  A CTRW is characterized by a distribution of hop lengths and dwell times, 
  where trajectories are characterized by sequential independent random draws from 
  each distribution.\cite{montroll_random_1965} FBM is common in crowded,
  viscoelastic environments where each jump comes from a Gaussian distribution
  but is anti-correlated to its previous step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
  An RWF is imposed by a system's geometry. Systems with tortuous pathways 
  and dead ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}

  We treated our system as an FBM process subordinate to a CTRW, or subordinated 
  FBM (sFBM) for short. We fit sFBM
  models in two ways: First, we parameterized solute motion using a single set of parameters fit
  to a hypothesized anomalous diffusion model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's radial
  distance from the closest pore center. This allows us to include different dynamical
  behavior of solutes observed within the pore region versus within the monomer tails.

@@

  Our second modeling approach applies an extended MSM framework to a relatively 
  small set of known states based on the three previously observed solute trapping mechanisms.  
  Standard MSMs are typically applied to determine equilibrium populations
  of states and the kinetics of transitions between those states.~\cite{bowman_using_2009} 
  We extend the framework to include state-dependent fluctuations and correlations in 
  solute position. The magnitude of solute fluctuations from its average trapped 
  position and the degree of correlation with previous fluctuations is 
  determined by the current state. To distinguish our approach from standard MSMs, we have
  named it the Markov state-dependent dynamical model (MSDDM).

  We determine the degree of success of our modeling approaches in two ways. First,
  if we can closely reproduce solute MSDs measured from MD simulations with realizations
  of our models, then it is likely that the model sufficiently captures solute 
  dynamics and can be used in a predictive capacity. Even if a model fails to 
  reproduce the MD MSDs, there is value in uncovering the cause of the deviation.
  The second measure of success is by analysis of the parameters defining each 
  model. The parameters themselves should provide clues that help explain solute 
  dynamics in a coherent way. A model that successfully reproduces MD MSDs, but 
  with nonsensical parameters has severely diminished value as it could be a 
  consequence of over-fitting.

  The goal of this work is not to definitively determine which modeling approach
  is better but to evaluate their performance independently because they both
  have potential value dependent on a given research study's goals. The anomalous subdiffusion
  approach provides a systematic way to compare the dynamical behavior of different
  solutes based solely on the time series of their center of mass positions. The process
  of choosing the correct anomalous subdiffusion provides its own mechanistic insight
  since it requires a thorough analysis of solute time series behavior. One can simulate
  different solutes, compare the fit model parameters and relate them back to 
  differences in solute size and chemical composition. 
  %MRS8: this may be a bit of a stretch, since it's not clear it helps that much in the end, but 
  %is fine to leave in now - maybe we revisit if needed during revisions.
  The MSDDM characterizes 
  explicitly defined trapping mechanisms, providing a clear picture of solute behavior
  while in each trapped state as well as the equilibrium occupation of each trapping state.
  It is possible to use the two models in tandem, the anomalous diffusion model to
  identify mechanisms, and the MSDDM to characterize mechanisms.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018~\cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.  %BJC4: version 2018. That's it not 2018.3 or anything. 
  We performed all post-simulation trajectory analysis using Python scripts which 
  are available online at \\ \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.
  The appropriate scripts to use for subsequent calculations are summarized in 
  Table~\ref{S-table:python_scripts} of the Supplemental Material.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an atomistic
  molecular model of four pores in a monoclinic unit cell with 10\% water by weight 
  (see Figure~\ref{fig:membrane_structure}). Approximately one third of the water 
  molecules occupy the tail region with the rest near the pore center. We chose to
  focus our study on the 10 wt\% water system because solutes move significantly 
  faster than in the 5 wt\% system studied previously thus allowing more statistically
  valid analyses.

@@

  In this study, we primarily use the MSD as a tool for characterizing the average
  dynamic behavior of solute trajectories. Rather than using them to calculate 
  diffusion constants or to relate our simulations to experimental measurement, we
  compare MSDs calculated from MD simulations to those generated from our models 
  in order to validate those models. Therefore, it is only important that we use a
  consistent definition for calculating the MSD between modeled trajectories and
  directly observed MD trajectories.

  One can measure MSD in two ways. The ensemble averaged MSD measures 
  displacements with respect to a particle's initial position:
  \begin{equation}
  \langle z^2(t) \rangle = \langle z(t) - z(0) \rangle^2
  \label{eqn:ensemble_msd}
  \end{equation}
  Fits to the \textit{ensemble} averaged MSD will always reproduce the form of 
  Equation~\ref{eqn:msd_form}. The time-averaged MSD measures all observed 
  displacements over time lag $\tau$: 
  \begin{equation}
  \overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory.

@@

  We chose to use just the time-averaged MSD to compare MD trajectory and modeled
  trajectories, because, compared to the ensemble average, it is a more 
  statistically robust measure of the average distance a solute travels over time.
  The ensemble MSD of only 24 solute trajectories would have much higher uncertainties.

@@

  \textit{Correlated hop length distributions:} The marginal distribution of hop 
  lengths by solutes undergoing an sFBM process is Gaussian, therefore we 
  parameterize it by its standard deviation, $\sigma$.
  ~\cite{metzler_random_2000, metzler_anomalous_2014,neusius_subdiffusion_2009}
  The measured mean of the hop length distribution is always very close to zero so
  we assume that it is exactly zero in our time series simulations since we have 
  no reason to expect drift in either direction based on pore symmetry.

@@

  Because we also want to account for the possibility that distribution of hops is 
  not Gaussian, we can model them with the more general class of L\'evy stable 
  distributions. For independent and identically distributed random variables, 
  the generalized central limit theorem assures convergence of the associated 
  probability distribution function (PDF) to a L\'evy stable PDF. \cite{klages_anomalous_2008}
  The characteristic equation which describes the Fourier transform of a L\'evy stable PDF is: 
  \begin{equation}
    p_{\alpha_h, \beta}(k;\mu,\sigma) =\exp\left[i\mu k - \sigma^{\alpha_h}|k|^{\alpha_h}\left(1 - i\beta\frac{k}{|k|}\omega(k, \alpha_h)\right)\right]
  \end{equation}
  where \\
  \[\omega(k, \alpha_h) = \begin{cases}
  	\tan{\frac{\pi \alpha_h}{2}} & \text{if}~\alpha_h \neq 1, 0 < \alpha_h < 2, \\
  	-\frac{2}{\pi}\ln |k| & \text{if}~\alpha_h = 1
  	 \end{cases}
  \]
  $\alpha_h$ is the index of stability or L\'evy index, $\beta$ is the skewness 
  parameter, $\mu$ is the shift parameter and $\sigma$ is a scale parameter. The most
  familiar case, and one of three that can be expressed in terms of elementary functions,
  is the Gaussian PDF ($\alpha_h$ = 2, $\beta$ = 0). We assume symmetric distributions
  centered about 0 implying that $\beta$ and $\mu$ are both 0.

@@

  We observe different dynamical behavior when solutes move while inside 
  the pore versus while in the tail region. This suggest two anomalous diffusion 
  models of varying complexity. We first create a simple, single mode model with
  a single set of parameters fit to solute trajectories.
  Our second, two mode model assigns a set of parameters to each of 2 modes based
  on the solute's radial location. We define the first mode as the pore region, 
  defined as less than 0.75 nm from any pore center. Solutes outside the pore 
  region are in the second mode, the tail region. We determined this cut-off by
  maximizing the difference in dynamical behavior as 
  described in our previous work.~\cite{coscia_chemically_2019} Unfortunately, 
  there were not enough sufficiently long sequences of hops in each mode to reliably
  calculate a Hurst parameter for each mode so we used the single, average Hurst 
  parameter from the single mode model for both modes of the two mode model.

@@

%BJC8: from author guidelines. Should there only be a line between column headings?
%Use a single horizontal rule to separate the column headings from the entries in the table. Use horizontal space within the body of the table to separate broad groups of entries. Extra vertical space may be used between columns. Avoid vertical rules in tables. Use lower-case roman letters for footnotes placed at the bottom of a table.

  \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ccc}
	  %\hline
	  Dwell Distribution                & ~~~~~~~~~Hop Length Distribution~~~~~~~~~~~ & Abbreviation \\
	  \hline
      Power Law                         & Gaussian                & sFBM         \\
      Power Law w/ Exponential Cut-off  & Gaussian                & sFBMcut      \\
      Power Law                         & L\'evy Stable           & sFLM         \\
      Power Law w/ Exponential Cut-off  & L\'evy Stable           & sFLMcut      \\
	  %\hline
	  \end{tabular}
	  \caption{We tested four anomalous diffusion models with various modifications 
	  to the dwell and hop length distributions. We incorporate correlation structure
	  into all models.}\label{table:anomalous_models}
 \end{table}

@@

  %BJC3: not sure how to cite fbm
  We used the Python package \texttt{fbm} to generate exact simulations of FBM and our
  own Python implementation of the algorithm by Stoev and Taqqu to simulate FLM.
  \cite{stoev_simulation_2004} Note that, to our knowledge, there are no known exact
  simulation algorithms for generating FLM trajectories. However, the algorithm
  we used sufficiently approximates draws from the marginal distribution and 
  reasonably approximates the correlation structure on MD simulation timescales. We
  added an empirical correction to enhance the accuracy of the correlation structure
  (see Section~\ref{S-section:flm_correlation} of the Supplemental Material for 
  validation of the approach).

@@

  A Markov state model (MSM) decomposes a time series into a set of discrete states
  with transitions between states defined by a transition probability matrix, $T$,
  which describes the conditional probability of moving to a specific state given
  the previously observed state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
  %MRS6: maybe skip this next sentence since you don't actually use them?
%  Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011} and 
%  pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable of featurization and
%  dimensional reduction which are often necessary to identify representative states.

  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. Therefore, there is no need to 
  apply any algorithmic approaches to identify and decompose our system into 
  discrete states. The states we have chosen include all combinations of trapping
  mechanisms in the pore and out of the pore (see Table~\ref{table:states}). They
  assume that there are no significant kinetic effects resulting from solute 
  conformational changes or pore size fluctuations. We use the same radial cut-off
  (0.75 nm) as in the AD model to differentiate the pore and tail region.  We define 
  a hydrogen bond to exist if the distance between donor, D, and acceptor, A, 
  atoms is less than 3.5 \AA~and the angle formed by $D-H \cdots A$ is less than
  30$\degree$.~\cite{luzar_effect_1996} We define a sodium ion to be associated 
  with an atom if they are within 2.5 \AA~of each other, as determined in our 
  previous work~\cite{coscia_chemically_2019}

  %BJC8: reformatted. I think this is what they want in terms of format
  \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ll}
	  \multicolumn{2}{c}{Markov State-Dependent Dynamical Model State Definitions} \\
	  \hline
	  1. In tails, not trapped                      & 5. In pores, not trapped                     \\
	  2. In tails and hydrogen bonding              & 6. In pores and hydrogen bonding             \\
	  3. In tails and associated with sodium        & 7. In pores and associated with sodium       \\
	  4. In tails, hydrogen bonding and associated  ~~~~& 8. In pores, hydrogen bonding and associated \\
	  %\hline
	  \end{tabular}
	  \caption{We defined 8 discrete states based on all combinations of previously observed solute
	  trapping mechanisms.}\label{table:states}  
  \end{table}

@@

  Adding to the standard MSM framework, we incorporated the dynamics of the solutes
  within each state as well as the dynamics of state transitions, which includes the
  overall configurational state of the solute and it's environments. 
  While MSMs are often used to estimate equilibrium populations of various states, adding 
  state-dependent dynamics allows us to simulate solute trajectories.
  %BJC5: I said the following earlier , but I still think it belongs
  Hence why we refer to them as Markov state-dependent dynamical models (MSDDMs).

@@

  We use the Hurst parameter to describe this negative time series correlation.
  However, there is not sufficient data to accurately measure a Hurst parameter for 
  each type of transition. We avoided this problem by combining all distributions 
  associated with state transitions and treating all transitions as correlated
  emissions from a single L\'evy stable distribution. This reduces the number of emission
  distributions from 64 to 9 (1 for each of the 8 states and 1 for transitions between states).

@@

  %In both case, selectivity is enhanced as the pore 
  %length increases which comes at the cost of lower solute flux, a phenomenon 
  %ubiquitous in polymer separations membranes.~\cite{geise_water_2011}

>>>>

  Mathematically modeling complex transport phenomena at the molecular level 
  can be a powerful tool for identifying transport mechanisms and predicting
  macroscopic properties. We use two different stochastic time series models,
  parameterized from long molecular dynamics (MD) simulation trajectories of
  a cross-linked H\textsubscript{II} phase lyotropic liquid crystal (LLC) 
  membrane, in order to predict solute mean squared displacements (MSDs) 
  and solute flux, and thus solute selectivity, in macroscopic length pores. 
  First, using anomalous diffusion theory, we show how solute dynamics can be
  modeled as a fractional diffusion process subordinate to a continuous time 
  random walk. From the MD simulations, we parameterize the distribution of 
  dwell times, hop lengths between dwells and correlation between hops. We 
  explore two variations of the anomalous diffusion modelling approach. The first variation
  applies a single set of parameters to the solute displacements and the second
  applies two sets of parameters based on the solute's radial distance from the
  closest pore center. Next, we generalize Markov state models, treating the
  configurational state of the system as a Markov process, each
  of which having separate transport properties. For each state and transition
  between states, we parameterize the distribution and temporal correlation 
  structure of positional fluctuations as a means of characterization and to 
  allow us to predict solute MSDs. We show that both models reasonably reproduce
  the MSDs calculated from MD simulations. However, qualitative differences 
  between MD and Markov state dependent model-generated trajectories may limit 
  its usefulness. Finally, we demonstrate how one can use these models to 
  estimate flux of a solute across a macroscopic-length pore and, based on 
  those quantities, the membrane's selectivity towards each solute. This work 
  helps to connect microscopic chemically-dependent solute motions that do not
  follow simple diffusive behavior with macroscopic membrane performance.

@@

  Lyotropic liquid crystals (LLCs) are a class of amphiphilic molecules that
  can be cross-linked into mechanically strong and highly selective nanoporous
  membranes.~\cite{gin_polymerized_2008} They may provide a promising
  alternative to conventional membrane separation techniques by being selective
  based not only on solute size and charge, but on solute chemical
  functionality as well.~\cite{dischinger_application_2017} One can tune the
  functionality of LLC monomers in order to enhance or weaken specific
  solute-membrane interactions.~\cite{dischinger_effect_2017} Two general
  morphologies of LLC membranes are being actively developed. The inverted
  hexagonal, H\textsubscript{II}, LLC phase has densely packed, uniform-sized
  pores on the order of 1 nm in size. A perfectly synthesized
  H\textsubscript{II} phase has an ideal geometry for high throughput
  separations. Unfortunately, aligning the microscopic hexagonal mesophases on
  a large scale as required for high permeability has shown limited
  success.~\cite{feng_scalable_2014,feng_thin_2016} Bicontinuous cubic,
  Q\textsubscript{I}, phase LLC membranes share the uniform and nm-sized chemically
  complex pores of the H\textsubscript{II} phase but its geometry consists of a
  tortuous network of 3D interconnected pores which do not require
  alignment.~\cite{carter_glycerol-based_2012} While its pore structure may
  decrease a Q\textsubscript{I} phase membrane's permeability relative to the
  H\textsubscript{II} phase, it is considerably easier to synthesize at
  scale.~\cite{hatakeyama_nanoporous_2010}

@@

  Using a bottom-up modeling approach, we can parameterize single particle
  behavior by extracting their trajectories from MD simulations and studying
  the properties that have the greatest influence on solute dynamics. With this
  information, we can learn how to construct an ensemble of characteristic
  single solute trajectories with these properties which would be useful for
  making long timescale predictions with computational ease and lower
  uncertainties. Fortunately, there is an abundance of information encoded in
  the complex solute trajectories. Into a model, we can incorporate fluctuations and
  time-dependent correlations in solute position as well as the length of time
  solutes are trapped. We can add further detail by integrating
  this time series analysis with our knowledge of the primary trapping
  mechanisms as well as the solute's changing chemical environment within the
  heterogeneous membrane.

  In this work, we use the output of our MD simulations to construct two
  classes of mathematical models which aim to predict membrane performance
  while providing quantitative mechanistic insights. The functional forms of
  these models are driven by mechanistic observations from our previous work and
  their inputs are parameterized using a substantial amount of data generated
  by long ($\geq 5~\mu$s) MD simulations.

@@

  Three well-known classes of behavior leading to anomalous subdiffusion are
  continuous time random walks (CTRWs), fractional Brownian motion (FBM) and
  random walks on fractals (RWFs).\cite{meroz_toolbox_2015} These types of
  motion are frequently used alone or in combination to describe single
  particle trajectories.~\cite{morrin_three_2018,metzler_anomalous_2014} A CTRW
  is characterized by a distribution of hop lengths and dwell times, where
  trajectories consist of sequential independent random draws from each
  distribution.\cite{montroll_random_1965} FBM is common in crowded,
  viscoelastic environments where each jump comes from a Gaussian distribution
  but is anti-correlated to its previous
  steps.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
  An RWF is imposed by a system's geometry. Systems with tortuous pathways and
  dead ends cause anti-correlated
  motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}

  We treated our system as an FBM process subordinate to a CTRW, or
  subordinated FBM (sFBM) for short. We fit sFBM models in two ways: First, we
  parameterized solute motion using a single set of parameters fit to a
  hypothesized anomalous diffusion (AD) model. Second, we used a two-state approach
  by extracting two sets of model parameters calculated based on a solute's
  radial distance from the closest pore center. This allows us to include
  different dynamical behavior of solutes observed within the pore region
  versus within the monomer tails.

@@

  Our second modeling approach applies an extended MSM framework to a
  relatively small set of known states based on the three previously observed
  solute trapping mechanisms. Standard MSMs are typically applied to determine
  equilibrium populations of states and the kinetics of transitions between
  those states.~\cite{bowman_using_2009} We extend the framework to include
  state-dependent fluctuations and correlations in solute position. The
  magnitude of a solute's fluctuations from its average trapped position and the
  degree of correlation with previous fluctuations is determined by the current
  state. To distinguish our approach from standard MSMs, we have named it the
  Markov state-dependent dynamical model (MSDDM).

  We determine the degree of success of our modeling approaches in two ways.
  First, if we can closely reproduce solute MSDs measured from MD simulations
  with realizations of our models, then it is likely that the model
  sufficiently captures solute dynamics and can be used in a predictive
  capacity. Even if a model fails to reproduce the MD MSDs, there is value in
  uncovering the cause of the deviation. 
  %BJC9: I modified the following from the commented out version below it in order to be more consistent with what we actually did.
  The second measure of success is based on the qualitative comparison between
  individual realizations of solute trajectories generated by our models and
  those observed from MD simulations. Even if we can reproduce the solute MSDs
  based on realizations of our models, the absence or inappropriate
  reproduction of hopping and trapping behavior may indicate underlying model
  issues. 
  %The second measure of success is by
  %analysis of the parameters defining each model. The parameters themselves
  %should provide clues that help explain solute dynamics in a coherent way. A
  %model that successfully reproduces MD MSDs, but with nonsensical parameters
  %has severely diminished value as it could be a consequence of over-fitting.

  The goal of this work is not to definitively determine which modeling
  approach is better but to evaluate their performance independently because
  they both have potential value dependent on a given research study's goals.
  The AD approach provides a systematic way to compare the dynamical behavior
  of different solutes based solely on the time series of their center of mass
  positions. The process of choosing the correct AD approach provides its own
  mechanistic insight since it requires a thorough analysis of solute time
  series behavior. One can simulate different solutes, compare the fit model
  parameters and relate them back to differences in solute size and chemical
  composition. The MSDDM characterizes explicitly defined trapping mechanisms,
  providing a clear picture of solute behavior while in each trapped state as
  well as the equilibrium occupation of each trapping state.  It is possible to
  use the two modelling approaches in tandem, the AD approach to identify
  mechanisms, and the MSDDM to characterize mechanisms.

@@

  We ran all MD simulations and energy minimizations using GROMACS
  2018~\cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.
  We performed all post-simulation trajectory analysis using Python scripts
  which are available online at \\
  \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.  The appropriate
  scripts to use for subsequent calculations are summarized in
  Table~\ref{S-table:python_scripts} of the Supplemental Material.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 10\%
  water by weight (see Figure~\ref{fig:membrane_structure}). Approximately one
  third of the water molecules occupy the tail region with the rest near the
  pore center. We chose to focus our study on the 10 wt\% water system because
  solutes move significantly faster than in the 5 wt\% system studied
  previously, thus allowing more statistically valid analyses.

@@

  In this study, we primarily use the MSD as a tool for characterizing the average
  dynamic behavior of solute trajectories. Rather than using them to calculate 
  diffusion constants or to relate our simulations to experimental measurements, we
  compare MSDs calculated from MD simulations to those generated from our models 
  in order to validate those models. Therefore, it is only important that we use a
  consistent definition for calculating the MSD between modeled trajectories and
  directly observed MD trajectories.

  One can measure MSD in two ways. The ensemble averaged MSD measures 
  displacements with respect to a particle's initial position:
  \begin{equation}
  \langle z^2(t) \rangle = \langle z(t) - z(0) \rangle^2
  \label{eqn:ensemble_msd}
  \end{equation}
  Fits to the ensemble averaged MSD will always reproduce the form of 
  Equation~\ref{eqn:msd_form}. The time-averaged MSD measures all observed 
  displacements over time lag $\tau$: 
  \begin{equation}
  \overline{z^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (z(t + \tau) - z(t))^2 dt
  \label{eqn:tamsd}
  \end{equation}
  where T is the length of the trajectory.

@@

  We chose to use just the time-averaged MSD to compare MD trajectories with
  modeled trajectories, because, compared to the ensemble average, it is a more
  statistically robust measure of the average distance a solute travels over
  time. The ensemble MSD of only 24 solute trajectories would have much higher
  uncertainties.

@@

  \textit{Correlated hop length distributions:} The distribution of hop 
  lengths by solutes undergoing an sFBM process is Gaussian, therefore we 
  parameterize it by its standard deviation, $\sigma$.
  ~\cite{metzler_random_2000, metzler_anomalous_2014,neusius_subdiffusion_2009}
  The measured mean of the hop length distribution is always very close to zero so
  we assume that it is exactly zero in our time series simulations since we have 
  no reason to expect drift in either direction based on pore symmetry.

@@

  Because we also want to account for the possibility that the distribution of
  hops is not Gaussian, we can model them with the more general class of L\'evy
  stable distributions. For independent and identically distributed random
  variables, the generalized central limit theorem assures convergence of the
  associated probability distribution function (PDF) to a L\'evy stable PDF.
  \cite{klages_anomalous_2008} The characteristic equation which describes the
  Fourier transform of a L\'evy stable PDF is: 
  \begin{equation}
    p_{\alpha_h, \beta}(k;\mu,\sigma) =\exp\left[i\mu k - \sigma^{\alpha_h}|k|^{\alpha_h}\left(1 - i\beta\frac{k}{|k|}\omega(k, \alpha_h)\right)\right]
  \end{equation}
  where \\
  \[\omega(k, \alpha_h) = \begin{cases}
  	\tan{\frac{\pi \alpha_h}{2}} & \text{if}~\alpha_h \neq 1, 0 < \alpha_h < 2, \\
  	-\frac{2}{\pi}\ln |k| & \text{if}~\alpha_h = 1
  	 \end{cases}
  \]
  $\alpha_h$ is the index of stability or L\'evy index, $\beta$ is the skewness 
  parameter, $\mu$ is the shift parameter and $\sigma$ is a scale parameter. The most
  familiar case, and one of three that can be expressed in terms of elementary functions,
  is the Gaussian PDF ($\alpha_h$ = 2, $\beta$ = 0). We assume symmetric distributions
  centered about 0 implying that $\beta$ and $\mu$ are both 0.

@@

  We observe different dynamical behavior when solutes move while inside the
  pore versus while in the tail region. This suggests two anomalous diffusion
  models of varying complexity. We first create a simple, single mode model
  with a single set of parameters fit to solute trajectories. Our second, two
  mode model assigns a set of parameters to each of 2 modes based on the
  solute's radial location. We define the first mode as the pore region,
  defined as less than 0.75 nm from any pore center. Solutes outside the pore
  region are in the second mode, the tail region. We determined this cut-off by
  maximizing the difference in dynamical behavior as described in our previous
  work.~\cite{coscia_chemically_2019} Unfortunately, there were not enough
  sufficiently long sequences of hops in each mode to reliably calculate a
  Hurst parameter for each mode so we used the single, average Hurst parameter
  from the single mode model for both modes of the two mode model.

@@

%BJC8: from author guidelines. Should there only be a line between column headings?
%Use a single horizontal rule to separate the column headings from the entries in the table. Use horizontal space within the body of the table to separate broad groups of entries. Extra vertical space may be used between columns. Avoid vertical rules in tables. Use lower-case roman letters for footnotes placed at the bottom of a table.
%MRS9: I'd go with what they said.  Looking at PRE, it looks like you can put double lines above and below the table, though - I added them 
  \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ccc}
            \hline
	  \hline
	  Dwell Distribution                & ~~~~~~~~~Hop Length Distribution~~~~~~~~~~~ & Abbreviation \\
	  \hline
      Power Law                         & Gaussian                & sFBM         \\
      Power Law w/ Exponential Cut-off  & Gaussian                & sFBMcut      \\
      Power Law                         & L\'evy Stable           & sFLM         \\
      Power Law w/ Exponential Cut-off  & L\'evy Stable           & sFLMcut      \\
	  \hline
          \hline
	  \end{tabular}
	  \caption{We tested four anomalous diffusion models with various modifications 
	  to the dwell and hop length distributions. We incorporate hop correlation 
	  into all models.}\label{table:anomalous_models}

 \end{table}

@@

  %BJC3: not sure how to cite fbm
  %TODO: cite github page for fbm
  We used the Python package \texttt{fbm} to generate exact simulations of FBM
  and our own Python implementation of the algorithm by Stoev and Taqqu to
  simulate FLM.~\cite{stoev_simulation_2004} Note that, to our knowledge, there
  are no known exact simulation algorithms for generating FLM trajectories.
  However, the algorithm we used sufficiently approximates draws from the
  marginal L\'evy stable distribution and reasonably approximates the
  correlation structure on MD simulation timescales. We added an empirical
  correction to enhance the accuracy of the correlation structure (see
  Section~\ref{S-section:flm_correlation} of the Supplemental Material for
  validation of the approach).

@@

  A Markov state model (MSM) decomposes a time series into a set of discrete
  states with transitions between states defined by a transition probability
  matrix, $T$, which describes the conditional probability of moving to a
  specific state given the previously observed
  state.~\cite{pande_everything_2010,wehmeyer_introduction_2018}
%  Software packages such as MSMbuilder~\cite{beauchamp_msmbuilder2:_2011} and 
%  pyEMMA~\cite{scherer_pyemma_2015} provide work flows capable of featurization and
%  dimensional reduction which are often necessary to identify representative states.

  In this work, we define a total of 8 discrete states based on the 3 trapping
  mechanisms observed in our previous work. Therefore, there is no need to
  apply any algorithmic approaches to identify and decompose our system into
  discrete states. The states we have chosen include all combinations of
  trapping mechanisms in the pore and out of the pore (see
  Table~\ref{table:states}). They assume that there are no significant kinetic
  effects resulting from solute conformational changes or pore size
  fluctuations. We use the same radial cut-off (0.75 nm) as in the AD approach to
  differentiate the pore and tail region. We define a hydrogen bond to exist
  if the distance between donor, D, and acceptor, A, atoms is less than 3.5
  \AA~and the angle formed by $D-H \cdots A$ is less than
  30$\degree$.~\cite{luzar_effect_1996} We define a sodium ion to be associated
  with an atom if they are within 2.5 \AA~of each other, as determined in our
  previous work~\cite{coscia_chemically_2019}

  \begin{table}[!htb]
	  \centering
	  \begin{tabular}{ll}
            \hline
            \hline
	  \multicolumn{2}{c}{Markov State-Dependent Dynamical Model State Definitions} \\
	  \hline
	  1. In tails, not trapped                      & 5. In pores, not trapped                     \\
	  2. In tails and hydrogen bonding              & 6. In pores and hydrogen bonding             \\
	  3. In tails and associated with sodium        & 7. In pores and associated with sodium       \\
	  4. In tails, hydrogen bonding and associated  ~~~~& 8. In pores, hydrogen bonding and associated \\
	  \hline
          \hline
	  \end{tabular}
	  \caption{We defined 8 discrete states based on all combinations of previously observed solute
	  trapping mechanisms.}\label{table:states}  
  \end{table}

@@

  Adding to the standard MSM framework, we incorporated the dynamics of the
  solutes within each state as well as the dynamics of state transitions, which
  includes the overall configurational state of the solute and it's
  environments. While MSMs are often used to estimate equilibrium populations
  of various states, adding state-dependent dynamics allows us to simulate
  solute trajectories. Hence why we refer to them as Markov state-dependent
  dynamical models (MSDDMs).

@@

  We use the Hurst parameter to describe negative time series correlation.
  However, there is not sufficient data to accurately measure a Hurst parameter
  for each type of transition. We avoided this problem by combining all
  distributions associated with state transitions and treating all transitions
  as correlated emissions from a single L\'evy stable distribution. This
  reduces the number of emission distributions from 64 to 9 (1 for each of the
  8 states and 1 for transitions between states).

@@
